Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-10-08 paper.tex
more fixes

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,\tilde{v}'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | \tilde{v}' \rangle
		 \langle \tilde{v}' | 0 \rangle 
		 \frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | \tilde{v}' \rangle \mel{\tilde{v}'}{Q_z}{0} \frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{\tilde{v}'} 
			\langle \tilde{v}' | 0 \rangle 
			\frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}.
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
Note that to allow closure over $\ket{n}$, we write $\Delta_{mn, gv} \approx \Delta_{m, g}$, which is allowed since the system is assumed to be sufficently detuned from $\ket{n}$. \cite{Silverstein2012}
Here, $\langle \tilde{a} | b \rangle$ and $\mel{\tilde{a}}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively.
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

@@

The signatures of vibronic coupling in the HDFG spectrum are made clear by using a simple model.
Using a two-well system described by $H = H_g |g) \left(g| + H_e |e\right) (e|$ where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ is the dimensionless momentum coordinate and $q = \sqrt{\frac{\omega}{\hbar}} Q$ is the dimensionless position coordinate, the A and B contributions can be evaluated using Franck-Condon factors and Herzberg-Teller integrals (\autoref{fig:doubres_spec}a).
For simplicity, we restrict consideration to $v^\prime \in \{0,1,2\}$ and take $\Delta = 0.5$.

@@

Compared to $\vec{\mu}$ and $\alpha_{ij}$, it is difficult to measure $\beta_{ijk}$ values through spontaneous hyper-Raman experiments.\cite{Kelley2010}
Only a few experimental determinations of $\beta_{ijk}$ for vibrational modes have been performed.\cite{Xu1997, Shoute2005, Kelley2010}
Methods reported in the literature to measure absolute hyper-Raman polarizabilities depend upon internal standards such as hyperpolarizabilities of dissolved samples or two-photon absorption cross sections, also difficult to measure. \cite{Kelley2010} 
In the method introduced by Levenson and Bloembergen (Bloembergen interferometry experiment), an internal standard interferes with the resonant lineshape of a $\chi^{(3)}$ process. \cite{Levenson1974_1, Levenson1974_2}
Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
Since the Bloembergen interferometry experiment only relies on the third order susceptibility for measured species (e.g., benzene or CaF$_2$),\cite{Levenson1974_2} it is possible to use quantitative four wave mixing spectroscopy to calculate $\beta_{ijk}$ values for infrared active vibrations.
It is thus useful to investigate how a treatment of orientational averaging can extract $\beta_{ijk}$ from the HDFG $\chi^{(3)}_{IJKL}$ expression.

Using \autoref{eq:nfgamma} and \autoref{sivebeta}, we can write
\begin{equation}\label{chi3}
		\chi^{(3)}_{IJKL} = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle_{IJKL} \rho^{(0)}_{gg}\\
\end{equation}
For simplicity, we take $\rho^{(0)}_{gg} = 1$.
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Kauranen1996}
Substituting in \autoref{SIVEselection}, we see
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta_{ijk}}{\partial Q_v} {\frac{\partial \mu_l}{\partial Q_v}} \right\rangle_{IJKL} = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} \chi^{(3)}_{IJKL}
\end{equation}
Since $\abs{\partial \mu_l / \partial Q}$ values can be extracted from FT-IR spectra and $\chi^{(3)}$ can be extracted through the Bloembergen interferometry experiment,\cite{Levenson1974_1, Levenson1974_2, RN459} HDFG can give quantitative information on the magnitude of $\frac{\partial \beta_{ijk}}{\partial Q}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values.

@@

\section{Appendix: Comparison of vSFG and two-beam HDFG output}\label{appendixA}
To motivate the application of HDFG spectroscopy, we perform a calculation to compare vibrationally resonant two-color HDFG and vSFG output, where absorption effects are neglected for simplicity.
vSFG is a $\chi^{(2)}$ technique, when only vibrationally resonant, whose output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
The output polarizations for two-beam vSFG and HDFG, respectively, under the electric dipole approximation, and assuming exponential dephasing, when vibrationally resonant are written as:
	 	\begin{subequations}
		 	\begin{equation}\label{PSFG}
		 		\begin{split}
		 		\abs{P^{(2)}_{\text{vSFG}}} &= \frac{N_\text{surf}}{\ell} \abs{ \frac{F_{\text{vSFG}} \langle \alpha_{ij}\mu_{k} \rangle }{2D' \varepsilon_0 \hbar i \Gamma} E(\omega_2)E(\omega_1) } 
		 		\end{split}
			 \end{equation}
		 	\begin{equation}
		 		\begin{split}
			 		\abs{P^{(3)}_{\text{HDFG}}} &= 6 N_\text{bulk} \abs{{ \frac{-F_{\text{HDFG}} \langle \beta_{ijk} \mu_{l} \rangle }{6D \varepsilon_0 \hbar i \Gamma} E{(\omega_2)}E(\omega_2)E(-\omega_1) }}, 
		 		\end{split}
			 \end{equation}
		 \end{subequations}
where N$_\text{bulk}$ is the bulk number density ($\sim$ 10$^{28}$ m$^{-3}$ for liquid H$_2$O at room temperature), N$_\text{surf}$ is the surface number density ($\sim$ 10$^{16}$ m$^{-2}$ for H$_2$O adsorbed at quartz at room temperature),\cite{Du1994} $D'$ is the SFG degeneracy factor, and $\Gamma$ is the vibrational dephasing rate.
For a two-beam experiment, $D' = 2$ and $D = 3$.
The HDFG output gains a multiplicative factor of 6 because six singly resonant HDFG processes contribute at pulse overlap.

>>>>

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,\tilde{v}'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | \tilde{v}' \rangle
		 \langle \tilde{v}' | 0 \rangle 
		 \frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | \tilde{v}' \rangle \mel{\tilde{v}'}{Q_z}{0} \frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{\tilde{v}'} 
			\langle \tilde{v}' | 0 \rangle 
			\frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}.
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
Note that to allow closure over $\ket{n}$, we write $\Delta_{mn, gv} \approx \Delta_{m, g}$, which is allowed since the system is taken to be sufficently detuned from $\ket{m,n}$. \cite{Silverstein2012}
Here, $\langle \tilde{a} | b \rangle$ and $\mel{\tilde{a}}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively.
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

@@

The signatures of vibronic coupling in the HDFG spectrum are made clear by using a simple model.
Using a two-well system described by $H = H_g |g) \left(g| + H_e |e\right) (e|$ where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ is the dimensionless momentum coordinate, $q = \sqrt{\frac{\omega}{\hbar}} Q$ is the dimensionless position coordinate, and $\Delta$ is the offset of the excited state surface relative to the ground state, the A and B contributions can be evaluated using Franck-Condon factors and Herzberg-Teller integrals (\autoref{fig:doubres_spec}a).
For simplicity, we restrict consideration to $v^\prime \in \{0,1,2\}$ and take $\Delta = 0.5$.

@@

Compared to $\vec{\mu}$ and $\alpha_{ij}$, it is difficult to measure $\beta_{ijk}$ values through spontaneous hyper-Raman experiments.\cite{Kelley2010}
Only a few experimental determinations of $\beta_{ijk}$ for vibrational modes have been performed.\cite{Xu1997, Shoute2005, Kelley2010}
Methods reported in the literature to measure absolute hyper-Raman polarizabilities depend upon external standards, such as hyperpolarizabilities or two-photon excited fluorescence of known samples, also difficult to measure. \cite{Kelley2010} 
In the method introduced by Levenson and Bloembergen (Bloembergen interferometry experiment), an internal standard interferes with the resonant lineshape of a $\chi^{(3)}$ process. \cite{Levenson1974_1, Levenson1974_2}
Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
Since the Bloembergen interferometry experiment only relies on the third order susceptibility for measured species (e.g., benzene or CaF$_2$),\cite{Levenson1974_2} it is possible to use quantitative four wave mixing spectroscopy to calculate $\beta_{ijk}$ values for infrared active vibrations.
It is thus useful to investigate how a treatment of orientational averaging can extract $\beta_{ijk}$ from the HDFG $\chi^{(3)}_{IJKL}$ expression.

Using \autoref{eq:nfgamma} and \autoref{sivebeta}, we can write
\begin{equation}\label{chi3}
		\chi^{(3)}_{IJKL} = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle_{IJKL} \rho^{(0)}_{gg}\\
\end{equation}
For simplicity, we take $\rho^{(0)}_{gg} = 1$.
Orientational averaging shows polarization schemes can isolate specific linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Kauranen1996}
Substituting in \autoref{SIVEselection}, we see
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta_{ijk}}{\partial Q_v} {\frac{\partial \mu_l}{\partial Q_v}} \right\rangle_{IJKL} = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} \chi^{(3)}_{IJKL}
\end{equation}
Since $\abs{\partial \mu_l / \partial Q}$ values can be extracted from FT-IR spectra and $\chi^{(3)}$ can be extracted through the Bloembergen interferometry experiment,\cite{Levenson1974_1, Levenson1974_2, RN459} HDFG can give quantitative information on the magnitude of $\frac{\partial \beta_{ijk}}{\partial Q}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values.

@@

\section{Appendix: Comparison of vSFG and two-beam HDFG output}\label{appendixA}
To motivate the application of HDFG spectroscopy, we perform a calculation to compare vibrationally resonant two-color HDFG and vSFG output, where absorption effects are neglected for simplicity.
vSFG is a $\chi^{(2)}$ technique, when only vibrationally resonant, whose output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
The output polarizations for two-beam vSFG and HDFG, respectively, under the electric dipole approximation, and assuming exponential dephasing, when vibrationally resonant are written as:
	 	\begin{subequations}
		 	\begin{equation}\label{PSFG}
		 		\begin{split}
		 		\abs{P^{(2)}_{\text{vSFG}}} &= \frac{N_\text{surf}}{\ell} \abs{ \frac{F_{\text{vSFG}} \langle \alpha_{ij}\mu_{k} \rangle }{2D' \varepsilon_0 \hbar i \Gamma} E(\omega_2)E(\omega_1) } 
		 		\end{split}
			 \end{equation}
		 	\begin{equation}
		 		\begin{split}
			 		\abs{P^{(3)}_{\text{HDFG}}} &= 6 N_\text{bulk} \abs{{ \frac{-F_{\text{HDFG}} \langle \beta_{ijk} \mu_{l} \rangle }{4D \varepsilon_0 \hbar i \Gamma} E{(\omega_2)}E(\omega_2)E(-\omega_1) }}, 
		 		\end{split}
			 \end{equation}
		 \end{subequations}
where N$_\text{bulk}$ is the bulk number density ($\sim$ 10$^{28}$ m$^{-3}$ for liquid H$_2$O at room temperature), N$_\text{surf}$ is the surface number density ($\sim$ 10$^{16}$ m$^{-2}$ for H$_2$O adsorbed at quartz at room temperature),\cite{Du1994} $D'$ is the SFG degeneracy factor, and $\Gamma$ is the vibrational dephasing rate.
For a two-beam experiment, $D' = 2$ and $D = 3$.
The HDFG output gains a multiplicative factor of 6 because six singly resonant HDFG processes contribute at pulse overlap.
